---
title: "Workshop 2"
output: html_document
---

```{r}
### Load packages
library(tidyverse) # Collection of all the good stuff like dplyr, ggplot2 ect.
library(magrittr) # For extra-piping operators (eg. %<>%)
```


```{r}
# load the dataset from GitHub - original source
penguins <- read_csv("https://sds-aau.github.io/SDS-master/M1/data/cities.csv")
to_get_region= read_csv("https://sds-aau.github.io/SDS-master/M1/data/cities.csv")
```
```{r}
penguins %>% head()
```

```{r}
penguins %>% glimpse()
```

```{r}
# drop all missing observations 
penguins %<>% drop_na()
```


## Brief EDA

```{r}
penguins %>% count(region)
```




### PRincipal Component Analysis

```{r}
library(FactoMineR)
library(factoextra)
```
```{r}
res_pca <- penguins %>% 
  column_to_rownames("place")%>%
  select_if(is_numeric) %>%
  PCA(scale.unit = TRUE, graph = FALSE)

##column_to_rownames
```
We only select the columns that are values, and scale the variables. 

```{r,fig.align='center'}
res_pca %>% 
  fviz_screeplot(addlabels = TRUE, 
                 ncp = 10, 
                 ggtheme = theme_gray())
```

We can see the albow formation that we use 2 dimensions 


```{r}
res_pca$eig %>% as_tibble()
```
We usually would stop when the eigenvalue is below 1, but this will be at 5. 

```{r}
res_pca %>% get_pca_var()
```

```{r}
res_pca$var$coord %>% 
  as_tibble() %>% 
  head()
```
```{r,,fig.width=15,fig.height=10,fig.align='center'}
res_pca %>%
  fviz_pca_biplot(alpha.ind = "cos2", #
                  col.ind = "contrib", #how much does this variable contribute red more contributive 
                  gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
                  geom = "point", 
                  ggtheme = theme_gray()) 
```
Test

```{r,,fig.width=15,fig.height=10,fig.align='center'}
res_pca %>%
  fviz_pca_biplot(alpha.ind = "cos2", #kinda fejled, so more visiable the futher away from the middle
                  gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
                  geom = "point", 
                  ggtheme = theme_gray(),
                  addEllipses = TRUE,
                  habillage = to_get_region %>% pull(region) %>% factor() ) 
```
##Clustering


##on the dimentionaly data

```{r}
res_pca %>%
  glimpse()
```


```{r}
#penguins[,"pca1"]= res_pca$ind$coord[,1]
#penguins[,"pca2"]= res_pca$ind$coord[,2]
```


### K-means


```{r}
# install.packages('uwot') # If necessary install
library(uwot)
```

```{r,fig.align='center'}
# We use the viz_nbclust() function of the factorextra package
penguins %>%
  select_if(is_numeric) %>% 
  scale() %>%
  fviz_nbclust(kmeans, method = "wss")  
```

```{r}
res_km <- penguins %>% 
  select_if(is_numeric) %>%
  scale() %>% 
  kmeans(centers = 2, nstart = 20)  
```

```{r}
res_km 
```


```{r,,fig.width=15,fig.height=10,fig.align='center'}
res_km %>% 
  fviz_cluster(data = penguins %>% select_if(is_numeric) ,
               ggtheme = theme_gray())  
```
Under the hood it does a PCA when we just do it on the data. 

To see the cluster assignement to each observation we can see in the cluster variable 

```{r}
res_km$cluster

```


Put the cluster variable into the data


Create table to show region in each cluster 







